---
title: "Module06GuidedQs"
author: "Diana McSpadden"
date: "10/1/2020"
output: html_document
---

# Stat 6021: Guided Question Set 6

For this question we will use the data set “cereals.txt”. The data set contains nutritional
information on 77 brands of cereal. For this question, we will focus on the variables y =
calories (number of calories per serving), x1 = sugars (grams of sugars per serving), and
mfr indicating the manufacturer of the cereal brand. The manufacturers are coded G =
General Mills, K = Kellogg’s, and other = other manufacturers.

y = calories
x1 = sugars (grams/serving)
mfr = manufacturer G = General Mills, K = Kelloggs, O = Other

```{r loadDataAndGivens}
dataCereal <- read.table("cereals.txt", header=TRUE ,sep="")
attach(dataCereal)

n = 77

mfr<-factor(mfr) # this asks R to see mfr as a factor/categorical
is.factor(mfr) 

contrasts(mfr) 
```

**G is our reference class**

## Question 1

Create a scatterplot of the number of calories per serving against the grams of sugars
per serving for the 77 brands. Overlay separate regression lines for each of the three
manufacturers.
```{r q1-1}
##consider each region a subset
dataG <- subset(dataCereal,mfr=="G") 
dataK <- subset(dataCereal,mfr=="K") 
dataOther <- subset(dataCereal,mfr=="other") 

##fit 3 separate regressions, one for each region
mod1 <- lm(calories~sugars,data=dataG)
mod2 <- lm(calories~sugars,data=dataK)
mod3 <- lm(calories~sugars,data=dataOther)

##create a scatterplot with different colors and symbols for each region
plot(sugars,calories, main="Calories vs Grams Sugar per Serving, by Manufacturer")

points(dataG$sugars,dataG$calories, pch=1, col="darkgreen") 
points(dataK$sugars,dataK$calories, pch=2, col="red") 
points(dataOther$sugars,dataOther$calories, pch=12, col="blue")

abline(mod1,lty=1, col="darkgreen")
abline(mod2,lty=2, col="red") 
abline(mod3,lty=3, col="blue")

legend("topleft", c("General Mills","Kellogg","Other"), lty=c(1,2,3), pch=c(1,2,12), col=c("darkgreen","red","blue")) 
```

Based only on the scatterplot and the regression lines, describe the relationship between the variables of interest.

**Answer**
For General Mills, the slope is pretty flat. It doesn't appear that grams sugar per serving affects the calories per serving that much.

For Kellogg and Other manufacturers, there is a positive relationship between grams sugar per serving and calories per serving, and the relationship between sugar and calories for Kellogg and Other appear identical.


## Question 2 Create a regression model with interactions, i.e.,
y = β0 + β1x1 + β2I1 + β3I2 + β4x1I1 + β5x1I2 + E,

where I1 and I2 are indicator variables where 
I1 = 1 for a General Mills cereal and 0 otherwise, and 
I2 = 1 for a Kellogg’s cereal and 0 otherwise. 

Write down the estimated regression equation for this model.

**Question 2 Work**

Looks like we want **Other to be our reference variable**.
```{r q2-1}
mfr<-relevel(mfr, ref = "other") # now other will be the 0-0 class
contrasts(mfr)
```
Need to create an interactions model

```{r q2-2}
modelInteractions <- lm(calories~sugars*mfr)
summary(modelInteractions)
```

**E(Y|x) = B0 + B1x1 + B2I1 + B3I2 + B4x1I1 + B5x1I2**

**E(Y|x) = 85.93 + (2.87)x1 + (17.07)I1 + (2.28)I2 + (-1.81)x1I1 + (-0.16)x1I2**

## 3. Carry out the relevant hypothesis test ...
to see if the interaction terms can be dropped. If they can be dropped, re-fit the regression model, and write down the estimated regression equation for the model without interactions.

**Question 3 Work**

Need to compare moduleInteractions to modelAdditiveOnly (which I need to create)

```{r q3-1}
modelAdditiveOnly <- lm(calories ~ sugars + mfr)
anova(modelAdditiveOnly,modelInteractions) # compare additive vs. interaction
```

**Question 3 Answer**

The p-value for the Hypothesis Test shows that we fail to reject the null hypothesis (that the B4 and B5 == 0), so we will use the additive only model and not the interaction model.

## 4. Assess if the regression assumptions are met, for the model you will recommend to use. 

Also, be sure to carry out Levene’s test of equality of variances since we have a categorical predictor.

**Question 4 Work**

Need to run the various plots for the additive only model

```{r q4-1, fig.width=12, fig.height=10}

par(mfrow=c(2,2)) # 2 rows, 2 columns

##residual plot of additive only model
plot(modelAdditiveOnly$fitted.values,modelAdditiveOnly$residuals,main="Residual plot")
abline(h=0,col="red")

##ACF plot of residuals of additive only model
acf(modelAdditiveOnly$residuals, main="ACF plot")

##QQ plot of residuals of additive only model
qqnorm(modelAdditiveOnly$residuals)
qqline(modelAdditiveOnly$residuals, col="red")

##additional assumption to check with categorical predictor. 
# Is the variance of the response variable constant between all classes of the categorical predictor?
boxplot(calories~mfr, main="Boxplot of Calories by Manufacturer")
```
**Regression Assumptions**

**Mean 0 of residuals**: This assumption appears to be met based on residual plot.

**Standard variance of residuals of total model**: This assumption appears to be met based on the residual plot.

**Uncorrelated residuals based on lags**: This assumption is **not** met, at 1 and 15 lags in values the residuals are correlated.

**Normal Distribution of errors**: This assumption is met based on Normal QQ plot.

**Constant Variance of the response variable for Manufacturers**: This assumption is **not** met based on the Boxplot. General Mills has almost no variance in the response variable. "Other" manufacturer has more variance than Kellogg.

???? **QUESTION on VARIANCE** I am not sure if the difference between other and Kellogg is enough to be concerned about ????


**Now Levene's** test

H0: Variance is equal for all manufactures
Ha: Variance is not equal for at least one manufacturer

```{r q4-2}
##perform levene's test. Null states the variances are equal for all classes. 
library(lawstat)
levene.test(calories,mfr)
```
This is really **unexpected** when compared to the boxplot. 

If we use a 95% confidence, Levene's test says that we fail to reject H0 and our model **does** meet the assumption that there is constant variance of calories for each manufacturer and our differences in variance are based on our sample. 


## 5. Conduct pairwise comparisons ..
for the difference in mean calories among all pairs of manufacturers for given values of grams of sugars, i.e.,

(a) General Mills and other manufacturers,
(b) Kellogg’s and other manufacturers,
(c) General Mills and Kellogg’s.

Be sure to contextually interpret the results of these hypothesis tests to someone who doesn’t know statistics.

```{r q5-1}
library(multcomp)
pairwise<-glht(modelAdditiveOnly, linfct = mcp(mfr= "Tukey")) # glht = general linear hypothesis test
summary(pairwise)

modelAdditiveOnly$coef
```
**Interpretations**

The model is: y = 88.56 + (2.41 * sugars) + (3.62 IF General Mills) + (1.89 IF Kellogg)

The regression functions are:

**Other** manufacturer: g = 88.56 + (2.41 * sugars) + 0 + 0 = y == 88.56 + (2.41 * x1)

**General Mills**: g = 88.56 + (2.41 * sugars) + (3.62) == (88.56 + 3.62) + (2.41 * x1)

**Kellogg**: g = 88.56 + (2.41 * sugars) + (1.89) == (88.56 + 1.89) + (2.41 * x1)

(a) General Mills and other manufacturers: General Mills contains 3.62 more mean calories per serving for the same grams of sugar per serving as "Other" manufacturers.

(b) Kellogg’s and other manufacturers: Kellogg contains 1.89 more mean calories per serving when holding the grams of sugar per serving constant when compared to "Other" manufactures.

``` {r q5-2}
3.62 - 1.89
```

(c) General Mills and Kellogg’s: General Mills contains 1.73 more mean calories per serving when holding grams of sugar per serving constant when compared to Kelloggs.








