---
title: "Module 07 Guided Questions"
author: "Diana McSpadden"
date: "10/10/2020"
output: html_document
---

# Stat 6021: Guided Question Set 7

## Name: H. Diana McSpadden
### UID: hdm5s
### 10.10.2020


## Question 1) From the tutorial, the model with the "best" BIC has x2; x7; x8 as predictors. 
We will continue to work with this regression model for the rest of this question.

## Question 2) The PRESS statistic can be used in model validation as well as a criteria for model selection. 
Unfortunately, the regsubsets() function from the leaps package does not compute the PRESS statistic. 

The PRESS statistic can be written as:

PRESS = SUM [y-i - yi-hat]^2 = SUM [e-i / 1 - hii]^2

where, hii denotes the ith diagonal from the hat matrix

hat matrix == X * ((X'X)^-1) * X'


Write a function that computes the PRESS statistic for a regression model. 
Hint: the diagonal elements from the hat matrix can be found using the lm.influence() function.

```{r q2-1}

data<-read.table("nfl.txt", header=TRUE, sep="")
attach(data)

bestBICModel <- lm(y~x2+x7+x8)

summary(bestBICModel)

```
```{r q2-2}
# https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/lm

calculatePRESS <- function(theModel) {
  
  sumPRESS = 0
  hatDiagonals <- lm.influence(theModel)$hat
  
  for (i in 1:length(hatDiagonals)){
    sumPRESS <- sumPRESS + ((theModel$residuals[i]) / (1 - hatDiagonals[i]))^2
  }
  
  return(sumPRESS)
}


valuePRESS <- calculatePRESS(bestBICModel)

print(paste("PRESS is: ", valuePRESS))


### or

press <- function(theModel) {
  
  pr <- theModel$residuals / (1-lm.influence(theModel)$hat)
  
  return (sum(pr^2))
}

valuePRESS2 <- press(bestBICModel)

print(paste("PRESS 2 is: ", valuePRESS2))

```

## Question 3) Using the function you wrote in part 2, 
calculate the PRESS statistic for your regression model from part 1. 
```{r q3-1}
valuePRESS <- calculatePRESS(bestBICModel, hatDiagonals)

print(paste("PRESS is: ", valuePRESS))
```

Calculate the R2-Prediction for this model, and compare this value with its R2. 
```{r q3-2}
# https://www.rdocumentation.org/packages/car/versions/3.0-10/topics/Anova

anova(bestBICModel)

str(anova(bestBICModel))

sum(anova(bestBICModel)[2])
```

```{r q3-3}

r2ForBestBICModel <- summary(bestBICModel)$r.squared
print(paste("R Squared is: ",r2ForBestBICModel))

#alternative calculation: R2p = 1 - (SSE-p) / SST
p <- length(bestBICModel$coefficients)

SSE <- sum(bestBICModel$residuals**2)
SST <- sum(anova(bestBICModel)[2]) # I think this sums the sum squared column from the anova table.

r2p <- 1 - ((SSE) / SST)
print(paste("R Squared p Calculated: ", r2p))


r2AdjForBestBICModel <- summary(bestBICModel)$adj.r.squared
print(paste("R Squared Adjusted is: ",r2AdjForBestBICModel))

#alternative calculation:  1 - (((n-1) / (n-p)) * (1 - R2p))
n = nrow(data)

r2pAdj <- 1 - (((n-1) / (n-p)) * (1 - r2p))
print(paste("R Squared p Adj Calculated: ", r2pAdj))

```


What comments can you make about the likely predictive performance of this model?

About 75%of the variance in y is predicted by the model containing x2, x7, x8: passing yards, percent rushing, and opponents' rushing yards.

Adjusted takes into account the number of predictors, regular R does not.

## Question 4) Delete half the observations (chosen at random), and refit the regression model. 

```{r q4-1}
# https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/sample
#

sampleHalfData <- data[sample(nrow(data), (nrow(data) / 2), replace = FALSE, prob = NULL),]
#sampleHalfData

detach(data) # detach the full data
attach(sampleHalfData) # attach the half data

refitBestBICModel = lm(y~x2+x7+x8)
summary(refitBestBICModel)

```

**Calculate the R2-Prediction for this model.**
```{r q4-2}

r2ForRefitBestBICModel <- summary(refitBestBICModel)$r.squared
print(paste("Refit R Squared is: ",r2ForRefitBestBICModel))

#alternative calculation: R2p = 1 - (SSE-p) / SST
p <- length(refitBestBICModel$coefficients)

refitSSE <- sum(refitBestBICModel$residuals**2)
refitSST <- sum(anova(refitBestBICModel)[2]) # I think this sums the sum squared column from the anova table.

refitR2p <- 1 - ((refitSSE) / refitSST)
print(paste("Refit R Squared p Calculated: ", refitR2p))

refitR2AdjForBestBICModel <- summary(refitBestBICModel)$adj.r.squared
print(paste("Refit R Squared Adjusted is: ",refitR2AdjForBestBICModel))

#alternative calculation:  1 - (((n-1) / (n-p)) * (1 - R2p))
n = nrow(sampleHalfData)

refitR2pAdj <- 1 - (((n-1) / (n-p)) * (1 - refitR2p))
print(paste("Refit R Squared p Adj Calculated: ", refitR2pAdj))
```

**How well does this model predict the number of games won?**

Only ~64% of the variance in games won is explained by the model containing x2, x7, x8: passing yards, percent rushing, and opponents' rushing yards. The significance of the model using half the dataset is considerably lower than when using the full data set.


## Question 5) 5. Based on the models you fitted from the previous 2 parts, ...
compare the standard errors of the regression coefficients for both models.

```{r q5-1}

print("Coefficients for model with all data points:")
bestBICModel$coefficients

print(" ")
print("Coefficients for model half the data points:")

refitBestBICModel$coefficients

```

```{r q5-2}

summary(bestBICModel)

print(" ")

summary(refitBestBICModel)

hatDiagonalsRefit <- lm.influence(refitBestBICModel)$hat

calculatePRESS(refitBestBICModel)

```

**Answer: ** The half data model has about twice the standard error for each parameter as the full data model. It could mean we are over-fitted. It definately doesn't work well with half the data.

## Question 6) Using both models, calculate the predicted number of wins for all the NFL teams.

Compare these models in terms of how well they predict the number of wins for all the NFL teams by computing the SSres for both models.


# NEED TO REDO - he wants us to use the model created by the half data, on the full data!!!!!

```{r q6-1}

SSEFull <- sum(bestBICModel$residuals**2)
SSTFull <- var(data$y) * (nrow(data)-1)

print(paste("Sum Squared Residuals For Full Data Model: ", SSEFull))
print(paste("Sum Squared Total For Full Data Model: ", SSTFull))

SSEHalf <- sum(refitBestBICModel$residuals**2)
SSTHalf <- var(sampleHalfData$y) * (nrow(sampleHalfData)-1)

print(paste("Sum Squared Residuals For Half Data Model: ", SSEHalf))
print(paste("Sum Squared Total For Half Data Model: ", SSTHalf))

predictFull <- predict(bestBICModel, data)

residualsFull <- y - predictFull

predictHalf <- predict(refitBestBICModel, data)

residualsHalf <- y - predictHalf


sum(residualsFull)^2

sum(residualsHalf)^2
```


