---
title: "MOdeul08HW"
author: "Diana McSpadden"
date: "10/13/2020"
output: html_document
---

# Stat 6021: Homework Set 8

## H. Diana McSpadden
## UID: hdm52
## Date: 10/22/2020

In this question, you will revisit the swiss data set that you worked on in Homeworks 4 and 5. 
The data set contains information regarding a standardized fertility measure and socio-economic indictors for each of the 47 French-speaking provinces of Switzerland around the year 1888. 

In Homework 5, you found that the model with just three predictors: Education, Catholic, and Infant Mortality was preferred to a model with all the predictors. Fit the model with the three predictors, and answer the following questions.
```{r prepareModel}
library(ggplot2)

attach(swiss)
#?swiss

swissModel <-lm(Fertility ~ Education + Catholic + Infant.Mortality)


```
# Question 1
## (a) Are there any observations that are outlying in the response variable? 
Be sure to show your work and explain how you arrived at your answer.

** Work on q1a**

First, I will use R to get the externally studentized residuals:
```{r q1a-1}

##residuals
#regResiduals <- swissModel$residuals # regular residuals

##studentized residuals
#studentResiduals <- rstandard(swissModel) # rstandard give studentized residuals == ri

##externally studentized residuals
extStudentResiduals <- rstudent(swissModel) # externally studentized 
```


Next, I will calculate the critical value to use for comparison. Externally studentized residuals follow a t distribution with n - p - 1 degrees freedom, so that is the critical value I will use.

``` {r q1a-2}
n<-nrow(swiss)
p<-4
##critical value using Bonferroni procedure
criticalValue <- qt(1-(0.05/(2*n)), n-p-1)
criticalValue

```

Next, I will sort the externally studentized residuals and plot with the +/- critical value also plotted. This will show if any of the absolute values of externally studentized residuals are greater than the critical value.

``` {r}
sort(extStudentResiduals)
plot.new()
plot(extStudentResiduals,main="Externally Studentized Residuals", ylim=c(-4,4))
abline(h=criticalValue, col="red")
abline(h=-criticalValue, col="red")

extStudentResiduals[abs(extStudentResiduals)>criticalValue]

```
**Answer Question 1a**: No, it does not appear that there are any outlying responses.


## (b) Are there any observations that have high leverage? 
Be sure to show your work and explain how you arrived at your answer.

**Work on q1b**

First, I will get the leverage values from that hat matrix. The diagonal of the **hat matrix** are the standardized measures of the distance of the i-th observation from the center of the x space. 

**Large hat diagonals reveal observations that are potentially influential.**

```{r q1b-1}
lev<-lm.influence(swissModel)$hat
```

It is obvious that the average size of hat diagonal is p/n.

Thus, if a diagonal is **greater than 2p/n** then we consider it remote enough to be a **leverage point**.

``` {r q1b-2}

comparisonValue <- 2*p/n
comparisonValue

sort(lev)
```

I notice there are two high leverage points: **19 and 45**

For fun, I will plot the leverages with a plotted line to identify the high leverage "level" and the points that exceed that level.

``` {r q1b-3}

plot(lev, main="Leverages", ylim=c(0,0.5))
abline(h=2*p/n, col="red")
```

## (c) Are there any influential observations based on DFFITs and Cook's Distance?

**Work on q1c**

First, calculate DFFITS to see how drastically the fitted value changes with and without the presence of a particular observation.

Compare DFFITS to 2 * sqrt(p/n).

```{r q1c-1}

DFFITS<-dffits(swissModel)
DFFITS[abs(DFFITS)>2*sqrt(p/n)] # how drastically fitted value changes with and without the presence of the observation
```
Three points significantly change the fitted values of the model if they are not (or are) included in the creation of the model: points 6, 37, and 47.


Second, calculate DFBETAS to see how the significantly the coefficients will change with and without observations.

DFBETAS is compared to 2 / sqrt(n)

``` {r}
#DFBETAS - 
DFBETAS<-dfbetas(swissModel) # measures how estimated coefficients change without presence of the observation
DFBETAS[abs(DFBETAS)>2/sqrt(n)]
# how to find the data point
DFBETAS # can find that value with eyeballing
```

**Observations With HIgh DFBETAS: 6, 19, 37, 42**

Third, calculate COOKS to determine how fitted values change for all values, not just the removed observation, when an observation is removed.

COOKS is compared to an F distribution with p, and n-p degress of freedom.

``` {r}
# COOKS considered how fitted values changes for all values (not just the observation) when data point is removed.
COOKS<-cooks.distance(swissModel)
COOKS[COOKS>qf(0.5,p,n-p)] # f distribution

```
**There are no observations that need to be considered based on COOK'S Distance.**

## (d) Briefly describe the difference in what DFFITS and Cook's distance are measuring.

DFFITs checks if the fitted value for the removed observation predictor value(s) significantly changes if the observation is removed.

Cook's distance checks the amount of change for all fitted values if an observation is removed from the data used to create the model.



