---
title: "Stat6021Module0"
author: "Diana McSpadden"
date: "8/22/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Module 0 Homework Exercises

### Question 1
Statistical theory tells us the distribution of the sample means with a xed sample
size, under certain circumstances. The sampling distribution is an approximation of
the density histogram of the sample means. We know the sample means vary from
sample to sample. The sampling distribution tells us the expected value (mean) of the
distribution, and the standard deviation of the sample means.
(a) Suppose the variable X follows a normal distribution with mean  and standard
deviation . Consider taking random samples, each with size n, repeatedly. What
is the sampling distribution of the sample mean, x?
(b) Suppose the variable X has an unknown distribution but known mean and
known standard deviation What is the name of the statistical theory that
informs us that the sampling distribution of the sample mean, x, can be well-
approximated by a normal distribution?

**Part a**
x-bar approx normal distribution mean mew, with sigma over square root n sd.

**Part b**
Central Limit Theorem

### Question 2
An automatic machine in a manufacturing process produces subcomponents. The
lengths of the subcomponents follow a normal distribution with a mean of 116 cm and
a standard deviation of 4.8 cm.
(a) Find the probability that one selected subcomponent is longer than 118cm.
(b) Find the probability that if 3 subcomponents are randomly selected, their mean
length exceeds 118cm.

```{r q2a}
1 - pnorm((118-116) / 4.8)
```
```{r q2b}
1 - pnorm((118-116) / (4.8 / sqrt(3)))

```
### Question 3

What are the goals of constructing a confidence interval?

**Answer**
1. Create an estimate of the unknown parameter of interest
1. Create a range of plausible (typically 95%) values for a parameter of interest
1. Provide a measure of uncertainty

### Question 4
How does increasing the confidence level affect the margin of error and the width of the confidence interval? Hint: sketching the standard normal distribution will be helpful.

** Answer **
Increasing the confidence level, from say 95% to 99% means that the width of the confidence interval also increases because we must include a larger area of values to be more confident. The margin of error also increased because the multiplier increases as alpha decreases.

### Question 5
How does increasing the sample size affect the margin of error and the width of the confidence interval? Briefy explain.

**Answer**
Increasing the sample size decreases the margin of error because sigma/sqrt(n) gets smaller as the denominator increases cuase the multiplier to be smaller. We get closer to a standard normal distribution of the sample mean.

### Question 6
Use R to find the value of the t-multiplier when constructing a confidence interval for the mean in the following situations:
(a) 94% confidence interval with n = 49.
(b) 86% confidence interval with n = 82.
(c) 74% confidence interval with n = 150.

```{r q6a}
qt(0.96,48)
```

```{r q6b}
qt(0.93,81)
```
```{r q6c}
qt(0.87,149)
```
### Question 7
A random sample of 100 students had a mean grade point average (GPA) of 3.2 with
a standard deviation of 0.2.
(a) Calculate a 97% confidence interval for the mean GPA for all students.
(b) What is the margin of error for the condence interval found in the previous part?
What is the margin of error telling us?
(c) Based on this condence interval, is it reasonable to say that the mean GPA of
all students is 3.25 or greater?

```{r q7a}
# CI == 3.2 +- tmultiplier

tmult = qt(0.985,99)
moe = tmult * (.2 / sqrt(100))
lci = 3.2 - moe
uci = 3.2 + moe
ci = cat("(",lci,",",uci,")")
ci
```
```{r}
moe
```
No, gpa mean of 3.25 or greater is not reasonable.
### Question 8
determine if a change in data or simply normal variation.

### Question 9
Hypothesis questions are about the POPULATION

### Question 10
For each of the situations, state the appropriate null and alternative hypotheses, in
symbols and in words. Sketch how you would nd the p-value based on the calculated
test statistic.
(a) David's car averages 29 miles per gallon on the highway. He just switched to a
new motor oil that is advertised as increasing gas mileage. He wants to investigate
if the advertisement is accurate.

**Answer**
null: mew highway car mileage is 29 mpg
Ha: mew highway car mileage greater than 29mpg

(b) The diameter of a spindle in a small motor is supposed to be 4 millimeters. If
the spindle is too small or too large, the motor will not function properly. The
manufacturer wants to investigate whether the mean diameter is moved away from
the target.

**Answer**
H0: mew spindle diameter == 4mm
Ha: mew spindle diameter != 4mm


(c) The average time in traffic between 2 points of a congested highway used to be 2
hours. The government invested money to improve travel times by building extra
lanes and overpasses. Citizens want to access if travel times have improved, on
average.

**Answer**
H0: mew time == 2 hours
Ha: mew time < 2 hours

### Question 11
To have more evidence against the null hypothesis, our test statistic should be larger
/ smaller (choose one) in magnitude. Brie
y explain.

**Answer**
LARGER: equation is xbar - null mew / s/sqrt(n). Test statistic gets larger as it deviates from null mew.

### Question 12
How does increasing the difference between the sample mean and the population mean
under the null hypothesis affect the test statistic and the evidence against the null
hypothesis?

**Answer**
See q11

### Question 13
How does increasing the sample size affect the test statistic and the evidence against
the null hypothesis?

**Answer**
Decreases the denominator thus making test statistic larger. more evidence against 
h0.

### Question 14
Use R to obtain the critical values of the following hypothesis tests:
(a) H0 : mew = 3:5;Ha : mew != 3:5, with alpha = 0:08 and n = 96.
```{r q14a}
critvalue = qt(.96,95)
critvalue

```

(b) H0 :  = 75;Ha :  < 75, with  = 0:12 and n = 43.
```{r q14b}
critvalue2 = -qt(.88,42)
critvalue2
```


(c) H0 :  = 10;Ha :  > 10, with  = 0:045 and n = 132.
```{r q14c}
critvalue3 = qt(.955,131)
critvalue3
```

Hint: you will need either the qt() or pt() function.

## Question 15
Use R to obtain the p-values of the following hypothesis tests:
(a) H0 : mew = 48;Ha : mew != 48, with t-stat = 2:14 and n = 50.
```{r q15a}
(1 - pt(2.14,49)) * 2
```


(b) H0 : mew = 3;Ha : mew > 3, with t-stat = 0:78 and n = 316.
```{r q15b}
1 - pt(.78,315)
```


(c) H0 : mew = 12;Ha : mew < 12, with t-stat = 1:57 and n = 34.
```{r q15c}
pt(1.57,33)
```
Hint: you will need either the qt() or pt() function.

## Question 16
The 10-year historical average yield of corn in the United States is 160 bushels per acre.
A survey of 50 farmers this year gives a sample mean yield of 158.4 bushels per acre,
with a standard deviation of 5 bushels per acre. Does this sample provide evidence
that the yield of corn has decreased from the 10-year historical average? Conduct an
appropriate hypothesis test.
(a) State the null and alternative hypotheses.
**Answer**
H0: mew (average yield of corn in US) = 160 b/a
Ha: mew (average yield of corn in US) < 160 b/a


(b) Calculate the test-statistic.
**Work**
s = 5 b/a
n = 50
x-bar = 158.4
```{r}
tstat = (158.4 - 160)/(5 / sqrt(50))
tstat
```


(c) Find the p-value and critical value.
```{r}
pvalue = pt(-2.2637,49)
print(pvalue)

critvalue4 = -qt(.95,49)
print(critvalue4)
```


(d) State a conclusion in context.
```{r}

```


(e) How would you interpret the calculated p-value?
```{r}

```

## Question 17
Obtain the critical value of a hypothesis test where H0 : mew = 145;Ha : mew != 145, with
significance level alpha = 0:02. Suppose the sample size is 50.

mew: 145
alpha: 0.02
use .99 because two sided test
n = 50
df = 49
```{r q17}
critvalue17 = qt(0.99, 49)
critvalue17
```
## Question 18
Obtain the t-multiplier for a 98% confidence interval. Suppose the sample size is 50.

CI == 2-sided test, so use .99 for perc
n = 50
df = 49

```{r q18}
tmult18 = qt(0.99, 49)
tmult18
```
## Question 19
Compare the critical value and the t-multiplier found in the previous 2 parts. What is
an implication based on this comparison?

**Answer**
Crit value and CI are same for given perc and n.

## Question 20
Suppose the hypothesis test in part 17 is carried out and the p-value is 0.043. Which
of the following confidence intervals is/are possible?

**Q 17 info**
mew: 145
alpha: 0.02
use .99 because two sided test
n = 50
df = 49

**Ranges**
- (143.2, 144.5)
- (151.3, 154.6)
- (144.5, 163.5)
```{r}
(143.2 - 144.5) / 2
```
```{r}
(151.3 - 154.6) / 2
```
```{r}
(144.5 - 163.5) / 2
```

 
**Reasoning**
possible would contain mew: 145. Only (144.5, 163.5) contains 145, but at alpha .02 and p-value == 0.043 means pvalue > alpha of 0.02. We do not reject H0.

## Question 21
A random sample of 85 banded archerfish were collected, and their lengths were mea-
sured and recorded. Their average length was 20cm with a standard deviation of 3cm.
(a) Construct a 95% confidence interval for the population mean length of banded
archerfish.
```{r q21a}
n = 85
df = 84
mew = 20
sd = 3

alpha = .05
alphahalf = .025
alphaperc = 1 - (.05/2)

# 20 +/- ME
# 20 +/- qt(0.975,84)(3/sqrt(85))
mem = 20 - (qt(alphaperc,84)*(3/sqrt(85)))
mep = 20 + (qt(alphaperc,84)*(3/sqrt(85)))

cat("(",mem,",",mep,")")
```


(b) Based on your confidence interval, is it plausible that the population mean length
of banded archerfish is 21cm? Briefy explain.
**Answer**
No, a pop mean of 21cm is not plauible as the value ies outside the 95% confidence interval.


(c) Suppose you conduct the following hypothesis test. H0 : mew = 21;Ha : mew != 21.
Without actually performing any additional calculations, what do you expect the
p-value of this hypothesis test will be? Briefly explain.
* greater than 0.05
* less than 0.05
**Answer**
less than 0.05. 


(d) Conduct the hypothesis test to verify your answer to the previous part.
xbar = 20
mew = 21
sd = 3
n = 85
df = 84
```{r q21d}
tvalue21 = (20 - 21)/(3/sqrt(85))

pvalue21 = pt(tvalue21, 84)
pvalue21
```

**Answer**
0.00143 is smaller than 0.05 thus we reject the null hypothesis.


