---
title: "Module 01 Guided Qs"
output: html_notebook
---

# STAT 6021: Guided Question Set 1
The data for this question is saved as bp.txt". The data set contains information on the
weight (in pounds) and systolic blood pressure (in millimeters of mercury, mmHg) of 26
randomly selected males in the age group 25-30.

```{r getdata}
#getwd()
data <- read.table('bp.txt', sep='\t', header = TRUE)
head(data,5)
```

## Question 1

Create a scatterplot for this data set. Be sure to label the axes and give an appropriate
title. Based on the appearance of the plot, does a simple linear regression appear
reasonable for the data?

```{r}
plot(x=data$weight, y=data$BP, main='Weight v BP', xlab = 'Weight (lbs)', ylab = 'Systolic BP (mmHg)')
```


*Answer*
*Yes*, it does seem reasonable that sys BP is simple linear function of weight.

## Question 2

Perform a simple linear regression for systolic blood pressure against weight. Record
the estimated equation for the linear regression, and state the estimated values of
the slope and intercept. Interpret the values of the estimated slope and intercept, in
context.

