---
title: 'Stat 6021: Homework Set 1, Name: H. Diana McSpadden, UID: hdm5s'
output:
  html_document:
    df_print: paged
---
'''
# Name: H. Diana McSpadden
# UID: hdm5s
# Assignment: Homework Set 1
# Question 1
(R required) We will use the dataset "copier.txt" for this question. 


The Tri-City Office
Equipment Corporation sells an imported copier on a franchise basis and performs
preventive maintenance and repair service on this copier. The data have been collected
from **45 recent calls on users** to perform routine preventive maintenance service; for
each call, Serviced is the number of copiers serviced and Minutes is the total number
of minutes spent by the service person.

**Knowns**
n = 45
Serviced == number of copy machines serviced on a call
Minutes == number minutes spent by service person on the call

## First, get the data
'''
```{r getdata}
#getwd()
head(data,5)
```
'''
## 1a: What is the response variable in this analysis? What is predictor in this analysis?

**Answer**
The response variable is **Minutes**.
The predictor variable is **Serviced**.
In other words, we believe the number of copiers serviced predicts the length, in minutes, of the service call.

## 1b: Produce a scatterplot of the two variables. 
How would you describe the relationship between the number of copiers serviced and the time spent by the service
person?

'''
```{r scatterplot}
#attach the data frame
attach(data)

#use plot(...) to create a scatterplot
plot(x=Serviced, y=Minutes, main='Plot Minutes Against Number Copiers Serviced', xlab = 'Number of Copiers Serviced', ylab = 'Service Time (minutes)')
```
'''
**Relationship**
There is a strong linear relationship with positive correlation. I do not see any clusters or outliers.

##1c: Use the lm() function to fit a linear regression for the two variables. 
Where are the values of B1, B0, R2, and sigma-hat2 for this linear regression?
'''

```{r q1c}
lmodel = lm(data)
summary(lmodel)
```
```{r q1csigma}
lmodel_sigma = sigma(lmodel)
lmodel_sigma
```
'''
B1: 15.0352
B0: -0.5802
R2: 0.9575
sigma-hat2: 8.914

##1d: Interpret the values of B1, B0 contextually. 
Does the value of B0 make sense in this context?

**Answer**
B1 Interpretation: The change in service call time, in minutes, will be +15.0352 minutes for each additional copier serviced, after the first copier.

B0 Interpretation: B0 tells us that a service call for 0 minutes would take negative 0.5802 minutes, which does not makes sense, and a service call for 0 copiers also does not make sense.

##1e: Use the anova() function to produce the ANOVA table for this linear regression.
What is the value of the ANOVA F statistic? What null and alternative hypotheses are being tested here? What is a relevant conclusion based on this ANOVA F statistic?
'''

```{r q1e}
anova(lmodel)
```
'''
ANOVA F STAT: 968.66

H0: B1 == 0, or that the slope of the simple linear regression equation equals 0.
Ha: B1 != 0, or that the slope of the simple linear regression equation does not equal 0.

The F statistic is large (given the data set) are tells us that we should reject the null hypothesis and that our linear model predicts service call minutes based on number of copiers serviced better than using the average number of service call minutes, or any other number within the range of service minutes.

'''