---
title: "Module 5 Homework"
author: "Diana McSpadden"
date: "9/24/2020"
output: html_document
---

# Stat 6021: Homework Set 5

## Name: H. Diana McSpadden
### UID: hdm5s
### 10.01.2020


## Question 1: For this first question, ...
you will continue to use the dataset swiss which you also used in the last homework. Load the data. For more information about the data set, type ?swiss. The goal of the data set was to assess how fertility rates in the Swiss (French-speaking) provinces relate to a number of demographic variables.

### Q1(a) In the previous homework, you fit a model with the fertility measure ...
as the response variable and used all the other variables as predictors. Now, consider a simpler model, using only the last three variables as predictors: Education, Catholic, and Infant.Mortality. Carry out an appropriate hypothesis test to assess which of these two models should be used. State the null and alternative hypotheses, find the relevant test statistic, p-value, and state a conclusion in context. (For practice, try to calculate the test statistic by hand.)

```{r q1a-1}
# load the data
#?swiss

attach(swiss)
```

**About Swiss**

Standardized fertility measure and socio-economic indicators for each of 47 French-speaking provinces of Switzerland at about 1888.

A data frame with 47 observations on 6 variables, each of which is in percent, i.e., in [0, 100].

[,1]	Fertility	Ig, ‘common standardized fertility measure’

[,2]	Agriculture	% of males involved in agriculture as occupation

[,3]	Examination	% draftees receiving highest mark on army examination

[,4]	Education	% education beyond primary school for draftees.

[,5]	Catholic	% ‘catholic’ (as opposed to ‘protestant’).

[,6]	Infant.Mortality	live births who live less than 1 year.

**Question 1a Work**

```{r q1a-2}
swissModel.Full <-lm(Fertility~Education + Catholic + Infant.Mortality + Agriculture + Examination) # I put agriculture and examination on the end
swissModel.Full # view the full model coefficients

swissModel.Reduced <-lm(Fertility ~ Education + Catholic + Infant.Mortality)
swissModel.Reduced # view the reduced coefficients
```
 I will run the anova for the reduced model vs. the full model, and also for the full model.

```{r q1a-3}

# run the anova for the reduced model vs full model
anova(swissModel.Reduced,swissModel.Full)

```

**Question 1a Answer**

Hypothesis Test:

H0: B's for Agriculture, Examination = 0 (i.e. we use the reduced model)

Ha: at least one of B's for Agriculture, Examination != 0 (i.e. support the full model)

The p value is much to high, 0.5628, so I fail to reject the null hypothesis, and determine that the reduced model is appropriate in comparison to the full model.

**Calculate the Test Statistic By Hand**


r == number of predictors to drop == 2

n - p; n number of observations, p == number of parameters == 47 - 4 == 43

F r,n-p is what we compare to.

```{r q1a-4}
print(paste("Test Statistic: ",qf(0.95,2,43)))
```
### Q1(b) For the model you decide to use from part 1a, assess if the regression assumptions are met.
```{r q1b-1, fig.width=10, fig.height=8}

par(mfrow=c(2,2)) # 2 rows, 2 columns - there will be three panes to look through

# Residual plot
plot(swissModel.Reduced$fitted.values,swissModel.Reduced$residuals, main="MLR: Residuals vs. Fitted Values")
abline(h=0,col="red")

# ACF Plot
acf(swissModel.Reduced$residuals, main="MLR: ACF of Residuals")

# QQ Plot of Residuals
qqnorm(swissModel.Reduced$residuals)
qqline(swissModel.Reduced$residuals, col="red")

## BoxCox
library(MASS)
boxcox(swissModel.Reduced, lambda = seq(-0.5, 2, 0.01))
```
**Question Answer 1b**

The BOXCOX contains lambda of 1, so the constant variance assumption appears to be met.

The Residual plot shows an approximate 0 mean of residuals.

The QQ plot shows two "lumps" that indicate some non-normality.

Most importantly, the ACF plot shows that the residuals are correlated at a lag of 1, which does not meet the regression assumptions.


## Question 2 (No R) The data for this question come from 113 hospitals...
The key response variable is **InfctRsk**, the risk that patients get an infection while staying at the hospital.
We will look at five predictors:

* x1: Stay. Average length of stay at hospital

* x2: Cultures. Average number of bacterial cultures per day at the hospital

* x3: Age. Average age of patients at hospital

* x4: Census. The average daily number of patients

* x5: Beds. The number of beds in the hospital

**R Output**

Call: **lm(formula = InfctRsk ~ Stay + Cultures + Age + Census + Beds)**

Coefficients:
                        Estimate Std. Error t value Pr(>|t|)

(Intercept) 0.2051282     1.2075929         0.170     0.8654

Stay        0.2055252     0.0660885         3.110     0.0024 **

Cultures    0.0590369     0.0103096         5.726     9.5e-08 ***

Age         0.0173637     0.0229966         0.755     0.4519

Census      0.0010306     0.0034942         0.295     0.7686

Beds        0.0004476     0.0026781         0.167     0.8676

Residual standard error: 0.9926 on 107 degrees of freedom

Multiple R-squared: 0.4765, Adjusted R-squared: 0.4521

F-statistic: 19.48 on 5 and 107 DF, p-value: 9.424e-14

**Analysis of Variance Table**

Response: InfctRsk

            Df        Sum Sq        Mean Sq         F value     Pr(>F)
            
Stay        1         57.305        57.305        58.1676       1.044e-11 ***

Cultures    1         33.397        33.397        33.8995       6.154e-08 ***

Age         1         0.136         0.136         0.1376        0.71144

Census      1         5.101         5.101         5.1781        0.02487 *

Beds        1         0.028         0.028         0.0279        0.86759

Residuals   107       105.413       0.985

Only use the provided R output to answer the rest of part 2.

### (a) Based on the t statistics, which predictors appear to be insignificant?

**Answer Question 2a**

Based on the t-values Age, Census, and Beds appears insignificant.


### (b) Based on your answer in part 2a, carry out the appropriate hypothesis test to
see if those predictors can be dropped from the multiple regression model. Show all steps, including your null and alternative hypotheses, the corresponding test statistic, p-value, critical value, and your conclusion in context.

**Work on Question 2b**

H0: all the coefficients for Age, Census, and Beds = 0

Ha: at least one of the coefficients for Age, Census, or Beds != 0.

**Use Following calculations**
r == number of predictors to drops

n - p; n number of observations, p == number of parameters

F r,n-p is what we compare to.

```{r q2c-1}
# calculate that F statistic using SSR's and SSE
F0 = ((0.136 + 5.101 + 0.028) / 3) / (105.413 / 107)
print(paste("F0 is: ", F0))

# create the critical value
criticalFValue = qf(0.95,3,107)
print(paste("The critical value is: ", criticalFValue))

# create the pvalue
pvalue = 1-pf(F0,3,107)
print(paste("The p value for the F statistic is: ", pvalue))
```

**Answer Question 2b**
The p-value for the F-statistic is 0.155 which syas we fail to reject H0. The simpler model is appropriate with Stay and Cultures predictor variables.

### (c) Suppose we want to decide between two potential models:

* Model 1: using x1; x2; x3; x4 as the predictors for InfctRsk

* Model 2: using x1; x2 as the predictors for InfctRsk

Carry out the appropriate hypothesis test to decide which of models 1 or 2 should be used. Be sure to show all steps in your hypothesis test.


