---
title: 'Module02GuidedQs'
author: "Diana McSpadden"
date: "9/5/2020"
output: html_document
---

#QUESTIONS
1. what should we be using to decide if an F value is significant?


# Stat 6021: Guided Question Set 2
We will continue to use the dataset "bp.txt" that we saw in the last module. The dataset
contains information on the weight (in pounds) and systolic blood pressure (in millimeters
mercury, mmHg) of 26 randomly selected males in the age group 25-30.

```{r getdata}
data <- read.table("bp.txt", header = TRUE, sep = "\t")
head(data,10)
#data

attach(data) # attach data so that we can refer directly to column headers
```


## 1a. Produce a plot of BP against weight. 
```{r q1a}
plot(x=weight, y=BP, main='Plot Systolic BP Against Weight', xlab = 'Weight (lbs)', ylab = 'Systolic BP (mmHg)')
```

## 1b. Write the estimated linear regression equation.

**Answer: ** E(y|x) = B0-hat + B1-hat(x)

## 2a. Use the function cor() to obtain the correlation between systolic blood pressure and weight. 

```{r q2a}
corvalue <- cor(x=weight,y=BP) # what statistical value is this?
print(corvalue)

#model <- lm(BP~weight)
#summary(model)
```
## 2b. Interpret this value contextually. How reliable is this interpretation?
** Want to talk about this **
There is an "above average" linear relationship between weight and BP. Not sure how to answer the "reliable" question.

Pair the correlation with how the graph looks. Need to think about both.

Closer to +/- 1 then closer the relation.

Correlation value can be high, BUT if the plot doesn't look right take that into consideration.

.77 with only 26 data points, looks pretty strong.

## 3a. What is the change in the predicted systolic blood pressure when weight increases by one pound? 

```{r q3a}
model <- lm(BP~weight)
model
```

**Answer: ** For every 1 lb increase of weight the model estimates a .4194 increase in systolic blood preassure, measured in mmHg.


## 3b. Also report the corresponding 95% confidence interval for the change in the predicted systolic blood pressure when weight increases by one pound.

```{r q3b}
summary(model)
```


This is a confidence interval for B1-hat.

B1-hat +- (t-alpha/2, (n-2))(se(B1-hat))

```{r q3bcalc}

b1hat <- .4194

tmulti <- qt(0.975,24) # 2-sided 95% t value with 24 df
tmulti

Sxx <- sum((weight - mean(weight))^2)
Sxx

seB1hat <- sqrt(sigma(model) / Sxx)
seB1hat

#correction <- tmulti * seB1hat
#correction

#minB1hat <- b1hat - correction
#maxB1hat <- b1hat + correction

#print(paste("Manual min B1-hat",minB1hat))
#print(paste("Manual max B1-hat",maxB1hat))

#print("CI B1-hat: ", minB1hat, " - ", maxB1hat)

# Can also use confint
confint(model, level = 0.95)

#print(paste("95% CI for B1-hat: ", B1CI))

```
**Answer **
We are 95% confident that repeated samples of 26 weights will produce a model with a B1-hat between 0.2746 and 0.5642

Which means a 95% confidence that the BP will increase between 0.2746 and 0.5642 with each one pound of weight increase for the population of adult males 25 - 30 years old. This CI is for the population.

## 4a. Conduct a hypothesis test to determine whether or not there is a linear association between weight and systolic blood pressure. 

** Hypothesis **
H0: B1 = 0
Ha: B1 != 0

** Use an F test, or a t-test
```{r q4a}
anova(model)
```

**F TEST**
The F value == 35.744, which seems large enough (given the data set) to conclude there is a linear association between weight and BP.

The p-value of the F value is .000004 which is small enough to also indicate there is a linear association between weight and BP.

## 4b. State the hypotheses, p-value, and conclusion in context.

**Hypothesis: ** H0: there is no linear relationship between weight and BP., 

The p-value of the hypothesis test is .000004, which is small enough to reject H0.

We can conclude there is a linear relationship between weight and BP.

**??** It seems like we want to say that we conclude that there is not not a linear relationship **??** Is that too pedantic?

## 5. Are your results from parts 3 and 4 consistent? Briey explain.

**Answer**

t value is 5.97 which is really

In part 3, the CI for 95% CI in slope was small, which makes me think our model is pretty good, and then the F value and p-value also make me think we have a linear relationship. So, yes, I **think** they are consistent.

## 6a. Estimate the mean systolic blood pressure for young males who weigh 200 pounds.
```{r q6a}
yhat6a <- 69.1044  + (0.4194 * 200)
yhat6a
```

## 6b. Also report the 95% confidence interval for the mean systolic blood pressure of young males who weigh 200 pounds.
```{r q6b}
newweight <- data.frame(weight=200)
predict.lm(model,newweight,level=0.95, interval="confidence")
```
## 7. Report the 95% prediction interval for the systolic blood pressure a young male who weighs 200 pounds.
```{r q7}
predict.lm(model,newweight,level=0.95, interval="prediction")

```
## 8. It has been suggested that if the predicted systolic blood pressure increases by more than 0.35 mmHg when weight increases by one pound, there is an increased risk of heart disease. 

**Is there significant evidence that males in the age group 25-30 are at an increased risk of heart disease?**

Conduct an appropriate hypothesis test. 

**What is the null and alternative hypotheses, test statistic, and conclusion?**

**Hypothesis Test**
H0: B1 = .35 # checking for a specific value of B1, not just B1 = 0
H0: B1 > .35 # one-sided test

Need to run the test.

**Test Statistic**
t0 = (B1-hat - .35) / se(B1-hat)



