---
title: 'Stat 6021: Homework Set 1, Name: H. Diana McSpadden, UID: hdm5s'
author: "Diana McSpadden"
date: "9/7/2020"
output:
  word_document: default
  html_document: default
---

# Name: H. Diana McSpadden

# UID: hdm5s

# Assignment: Homework Set 2

### Worked with Chelsea AAlvarado, David Fuentes, Caprill Wright, and Abby Bernhardt

# Question 1
(R required) For this question, you will use the dataset "Copier.txt" for this question. This is the same data set that you used in the last homework. The Tri-City Office Equipment Corporation sells an imported copier on a franchise basis and performs preventive maintenance and repair service on this copier. The data have been collected from 45 recent calls on users to perform routine preventive maintenance service; for each call, Serviced is the number of copiers serviced and Minutes is the total number of minutes spent by the service person.

It is hypothesized that the total time spent by the service person can be predicted using the number of copiers serviced. Fit an appropriate linear regression and answer the following questions:

**Knowns**
n = 45
Serviced == number of copy machines serviced on a call
Minutes == number minutes spent by service person on the call

## 1a. Produce an appropriate scatterplot and comment on the relationship between the total time spent by the service person and the number of copiers serviced.

```{r q1a}
#getwd()
copierData <- read.table('copier.txt', sep='\t', header = TRUE)
head(copierData,5)
attach(copierData)

#use plot(...) to create a scatterplot
plot(x=Serviced, y=Minutes, main='Plot Minutes Against Number Copiers Serviced', xlab = 'Number of Copiers Serviced', ylab = 'Service Time (minutes)')
```

## 1b. What is the correlation between the total time spent by the service person and the number of copiers serviced? Interpret this correlation contextually.

```{r q1b}
corvalue <- cor(x=Serviced,y=Minutes) 
print(paste("Correlation between Service Time and Number of Copiers: ", corvalue))
```
Because the correlation value is so close to 1, i.e. 0.9785, and the plot supports a positive linear relationship between service time and copiers serviced, I conclude there is a linear relationship between the predictor variable, number of copiers serviced, and the response variable, service time.

## 1c. Can the correlation found in part 1b be interpreted reliably? Briey explain.

Yes, it can be reliably interpreted as a number after also analyzing the scatter plot to confirm a linear relationship. Both confirm a linear relationship. The correlation is not a reliable predictor unless both the scatter plot and correlation value are reviewed.

## 1d. Obtain the 95% confidence interval for the slope, B1.

```{r q1d}

# create a linear model
copierModel <- lm(Minutes~Serviced)
copierModel

#summary(copierModel)

# get the 95% CI for the Serviced coef
CIB1 = confint(copierModel, "Serviced", level = 0.95)

print(paste("95% CI of slope: ", CIB1[1], " - ", CIB1[2]))

```

## 1e. Suppose a service person is sent to service 5 copiers. Obtain an appropriate 95% interval that predicts the total service time spent by the service person.
```{r q1e}

## 95% CI for response when x=5
newdata<-data.frame(Serviced=5)
#print(newdata) # print newdata to understand what we get back

prediction = predict.lm(copierModel,newdata,level=0.95, interval="prediction")
#prediction

print(paste("The 95% CI for service minutes for a service call for 5 copiers is: ", prediction[2], " - ", prediction[3], " minutes"))

```

## 1f. What is the value of the residual for the first observation? Interpret this value contextually.

```{r q1f}
# get all the residuals from the model
copierResiduals <- copierModel$residual
#copierResiduals

# get the first residual
print(paste("The value of the first residual is: ", copierResiduals[1]))


```
**1f Interpretation**
The residual is the difference between the observed first data point, and the model's prediction for the first data point. So, the difference between y0 for the first x value, and y-hat for the first x value:

Residual1 = (y1 - Y1-hat)

So, the first value in our data set is **9.49 fewer minutes** than E(Y|x) for a service call for 2 copiers (the first x data point).

## 1g. What is the average value of the all the residuals? Is this value surprising (or not)? Briefly explain.

```{r q1g}

meanCopierResidual = mean(copierModel$residual)

print(paste("The mean of the residuals is: ", meanCopierResidual, " minutes."))

```

**1g Explanation**

No, this is not surprising. The mean of the residuals is 0, and -0.00000000000000026122 is basically 0 with some expected rounding error. Last week we proved mathematically that the sum of the residuals is 0, so the mean (sum(residuals)/n) == (0/n) == 0. This is the result of the definition of BLUE.


# Question 2 (No R required) 

A substance used in biological and medical research is shipped by airfreight to users in cartons of 1000 ampules. 
The data consist of 10 shipments. The variables are number of times the carton was transferred from one aircraft to another during the shipment route (transfer), and the number of ampules found to be broken upon arrival (broken). 

We want to fit a simple linear regression. A simple linear regression model is fitted using R. The corresponding output from R is shown next, with some values missing.


Call:
lm(formula = broken ~ transfer)

**Coefficients**:
Estimate                Std. Error      t value         Pr(>|t|)
(Intercept)     10.2000 0.6633          ______          ________ ***
transfer        4.0000 0.4690           ______          ________ ***

Residual standard error: 1.483 on 8 degrees of freedom

**Analysis of Variance Table**
Response: broken
                Df      Sum Sq  Mean Sq         F value         Pr(>F)
transfer        1       160.0   160.0           ______          ________ ***
Residuals       8       17.6    2.2

**Additional Knowns**

Predictor Variable, x: Number of transfers

Response variable, y: Number of broken ampules

n = 10

x-bar = 1

Sxx, sum(((x-i) - (x-bar)^2) == 10

## 2a. Carry out a hypothesis test to assess if there is a linear relationship between the variables of interest.

H0: B1 = 0

Ha: B1 != 0

Our estimated equation is: E(y|x) = 10.2 + 4(x)

The equation has B1-hat == 4.

F0 = 160 / 2.2 = 72.7272, this is a pretty large number, and implies a t-value = sqrt(72.7272) = 8.528

Using the t-value comparison, 8.528 is greater than 2.306 so we rejecting H0, and determine there is a linear relationship.


**This is evidence for a linear relationship between the variables.**


## 2b. Calculate a 95% confidence interval that estimates the unknown value of the population slope.

B1-hat - t-alpha/2,n-2(se(B1-hat)) <= B1-hat <= B1-hat + t-alpha/2,n-2(se(B1-hat))

4 - t-(.975,8)(se(B1-hat)) <= B1-hat <= 4 + t(.975,8)(se(B1-hat)) 

4 - (2.306)(se(B1-hat)) <= B1-hat <=  4 + (2.306)(se(B1-hat)) # t-value from textbook for 0.025 and 8 df


se(B1-hat) = 0.4690 

se(B1-hat) = 4 / 4.6904

So,
4 - (2.306)(0.4690) <= B1-hat <=  4 + (2.306)(0.4690)

4 - 1.0815 <= B1-hat <= 4 + 1.0815

**95% CI for population slope is 2.918 - 5.0814**


## 2c. A consultant believes the mean number of broken ampules when no transfers are made is different from 9. 

Conduct an appropriate hypothesis test (state the hypotheses statements, calculate the test statistic, and write the corresponding conclusion in context, in response to his belief).

H0: B0 = 9

Ha: B0 != 9

t0 = (B0-hat - 9) / se(B0-hat)

se(B0-hat) == sqrt((MS-res)(1/n + ((x-bar)^2/Sxx))

t0 = (10.2 - 9) / sqrt((2.2) * (.1 + .1)) # MS-res of 2.2 comes from ANOVA table

t0 = (10.2 - 9) / sqrt(.44)

t0 = 1.2 / .6633

t0 = 1.8091 < 2.306004


This test statistic is smaller than t(0.975,8), a 95% confidence level which is 2.306, so we do not reject the null hypothesis. There is not evidence to suggest the intercept does not equal 9.

##2d. Calculate a 95% confidence interval for the mean number of broken ampules and a 95% prediction interval for the number of broken ampules when the number of transfers is 2.

**MEAN CI: CI for mean number of broken ampules**

E(y|2) +- t(.0975,8)(sqrt((MS-res)(1/n + (2 - 1)/Sxx))

(10.2 + (4*2)) +- (2.306)sqrt((2.2)(.1 + ((2 - 1)^2/10)) # t value for 0.025 with df = 8 is from textbook, and MS-res is 2.2
 
18.2 +- (2.306) * sqrt((2.2)(.1 + .1))
 
18.2 +- (2.306) * sqrt(2.2 * .2)
 
18.2 +- (2.306) * (0.6633)
 
18.2 +- 1.53
 
**95% CI for mean ampules broken when there are 2 transfers is 16.67 - 19.73**


**PREDICTED VALUE CI: CI for predicted number of broken ampules when number of transfers is 2**

10.2 + 4(2) +- (2.306004)sqrt((2.2)(1 + .1 + ((2 - 1)^2/10))

18.2 +- (2.306004)(sqrt((2.2 * 1.2)))

18.2 +- (2.306004)(1.6248)

18.2 +- 3.7468

**95% CI for the predicted number of ampules broken when there are 2 transfers is 14.4532 - 21.9468**


## 2e. What happens to the intervals from the previous part when the number of transfers is 1? (Describe what happens without calculating)

The range of the two confidence intervals will be smaller, and I can describe why both logically and mathematically:

**Logically: ** 1 transfer is the mean number of transfers, so our predictions will be better, thus a smaller interval, because our predictive power is better at the mean of x.

**Mathematically: ** The correction (+/-) to y0 will be smaller because the numerator to part of the correction is (x0 - (x-bar)), which will be 1 - 1 == 0.


## 2f. What is the value of the F statistic for the ANOVA table?

F0 = 160 / 2.2 = **72.7272**


##2g. Calculate the value of R2, and interpret this value in context.

The R2, Coef of determination, formula is: SS-R / SS-T, i.e. the regression error divided by the total error and can be used as a measure of how well variations in x explain variations in y.

R2 = 160 / (160 + 17.6)

R2 = 160 / 177.6 == **.9009**

This value is close to 1, and implies that our model does demonstrate a linear relationship between number of transfers and ampules broken. R2 can be unreliable without seeing the data and a scatter plot of the data.


# Question 3 (No R required) 
**Suppose that the population slope for a straight-line relationship between y and x is 0.**

## 3a. Describe how the straight line would look in a plot of y versus x.

**Answer**
The line would be a straight, perfectly horizontal line at some unknown y-intercept. 


## 3b. Explain why a slope that is equal to 0 would indicate that y and x are not linearly related, and why a slope that is not equal to 0 would indicate that y and x are linearly related.

**Answer**

A slope of zero describes a y value that does not change, no matter the value of x. x could be anything, and y remains stable. Y is independent of x. An example is the speed of light, no matter what the hair color of a person is, the speed of light remains the same. That relationship has a slope of zero.

If the slope is not equal to 0, then the value of y does change, to some degree, as the value of x changes. As the change in x increases/decreases from the mean of x, the change in y will also increase/decrease from y's mean. That change expresses some type of linear relationship with a non-0 slope.
