---
title: "Module 09 HW"
author: "Diana McSpadden"
date: "10/27/2020"
output: html_document
---

# Stat 6021: Homework Set 9

## H. Diana McSpadden
## UID: hdm52
## Date: 10/30/2020

**Attended group with**: 


# Question 1. In an experiment testing the effect of a toxic substance, ...
1,500 insects were divided at random into six groups of 250 each. 

The insects in each group were exposed to a fixed dose of the toxic substance. A day later, the insects were observed. Death from exposure was recorded as 1, and survival 0. 

The file “insects.txt” contains the data from the experiment. 

* The first column denotes the dose level on a logscale, x; 
* the 2nd column denotes the sample size for each group, ni
* ; and the 3rd column denotes the number of insects that died in the group, yi

You will probably want to read the data in and give appropriate names for the columns.
```{r q1-setup1}

library(dplyr)

insectsData <- read.table("insects.txt", header=FALSE, sep="")

```


``` {r q1-setup2}

insectsData <- insectsData %>% rename(logdose = V1, size = V2, died = V3)

#names(insectsData)
```

```{r q1-setup3}

head(insectsData)
attach(insectsData)

#knowns
n <- 1500
groupn <- 250
groups  <- 6

```

## Q1 (a) Plot the sample log odds against the log dose, x. 
Does this plot suggest fitting a logistic regression model to be appropriate?
```{r q1a-1}
prop<-died/size

plot(logdose, log(prop/(1-prop)), xlab= "Log Dose", ylab="sample log odds")

```

**Answer Q1a**: Yes, this plot looks perfet for a logistic regression model because the relationship is almost perfectly linear.

## Q1 (b) Use R to fit the logistic regression model, and write the estimated logistic regression equation.

```{r q1b-1}
modelInsects <- glm(prop~logdose, family="binomial", weights=size) # fit model. The weight is the sample/group size
summary(modelInsects)

```




## Q1 (c) Interpret the estimated coefficient βˆ1 in context.


## Q1 (d) What are the estimated odds of death at log dose level x = 2?


## Q1 (e) Predict the probability of death from exposure at log dose level x = 2.


## Q1 (f) Carry out both the Pearson’s χ2 and deviance goodness of fit tests to check the fit ...
of this logistic regression model. Clearly state the null and alternative hypotheses,
test statistic, and conclusion.





# Question 2. (No R needed) A health clinic sent ...
fliers to its clients to encourage everyone to get a flu shot. In a follow-up study, 159 elderly clients were randomly selected and asked if
they received a flu shot. 

A client who received a flu shot was coded y = 1, and a client who did not receive a flu shot was coded y = 0. 

Data were also collected on their age,

* x1, health awareness rating on a 0-100 scale (higher values indicate greater awareness),
* x2, and gender, 
* x3, where males were coded x3 = 1 and females were coded x3 = 0. 

A first order logistic regression model was fitted and the output is displayed below. ...

## Q2 (a) Interpret the estimated coecient for x3, gender, in context.


## Q2 (b) Conduct the Wald test for B3. ...
State the null and alternative hypotheses, calculate the test statistic, and make a conclusion in context.


## Q2 (c) Calculate a 95% confidence interval for B3, and interpret the interval in context.


## Q2 (d) Comment on whether your conclusions from parts 2b and 2c are consistent.


## Q2 (e) Suppose you want to drop the coefficients for age and gender, B1 and B3. 
A logistic regression model for just awareness was fitted, and the output is shown below. ...


Carry out the appropriate hypothesis test to see if the coefficients for age and gender can be dropped.



## Q2 (f) Based on your conclusion in question 2e, what are the estimated odds of a client receiving the ...
flu shot if the client is 70 years old, has a health awareness rating of 65, and is male? 

What is the estimated probability of this client receiving the flu shot?


## Q2 (g) Based on your conclusion in question 2e, ...
interpret the estimated coefficient for age, in context. 

What does this interpretation tell us about the estimated probability of a client receiving the flu shot if the client is 71 years old, has a health awareness rating of 65, and is male, compared to the estimated probability found for a similar client who is a year younger?




