---
title: "Module09GuidedQs"
author: "Diana McSpadden"
date: "10/27/2020"
output: html_document
---
# Stat 6021: Guided Question Set 9


The Western Collaborative Group Study (WCGS) is one of the earliest studies regarding
heart disease. Data were collected from 3154 middle-aged males in California. Download
the file "wcsg.csv" and load it into R. We will focus on predicting the likelihood of getting
a heart attack based on the following predictors:

* age. Age in years
* sbp. Systolic blood pressure in mm Hg
* dbp. Diastolic blood pressure in mm Hg
* ncigs. Number of cigarettes smoked per day, on average.

The response variable is **chd69**, with a `1' indicating the person developed coronary heart
disease, and a `0' indicating the person did not develop coronary heart disease.
```{r}
# load the data

data<-read.csv("wcgs.csv", header=TRUE)
attach(data)
```

## Question 1
Before fitting a model, create some graphical summaries to see if there is a difference in
the distributions of the predictors among people who did and did not develop coronoray
heart disease. To create a boxplot, type boxplot(age~chd69) to see the distribution
of the variable age differs between people who did and did not develop coronary heart
disease. Create similar boxplots for the other variables. Which variables seem to
differentiate those who developed and did not develop coronary heart disease?

```{r}

boxplot(age~chd69)

boxplot(sbp~chd69)

boxplot(dbp~chd69)

boxplot(ncigs~chd69)
```
From the box plots age and ncigs seem to matter.

## 2. Use R to fit the logistic regression model using all the predictors listed above, and write the estimated logistic regression equation.

```{r q2-1}

result<-glm(chd69 ~ age + ncigs + sbp + dbp, family = "binomial")
summary(result)

# knowns
n <- 3154
p <- 5

```

E(y) = exp(-9.12 + (0.0666 * age) + (0.0242 * ncigs) + (0.019 * sbp) + (0.008 * dbp)) / ( 1 + exp(-9.12 + (0.0666 * age) + (0.0242 * ncigs) + (0.019 * sbp) + (0.008 * dbp)))

Try to write in the log pi / (1 - pi) = -9.12 + (0.0666 * age) + (0.0242 * ncigs) + (0.019 * sbp) + (0.008 * dbp)

```{r q2-2}
exp(0.0242)
```

## 3. Interpret the estimated coefficient for ncigs in context.

Controlling for age, sbp, and dbp The predicted odds ratio of a middle aged male in CA being Dxed with CHD for a one unit increase in number of cigarettes is 1.0245.

## 4. What are the estimated odds of developing heart disease for an adult male who is ..
45 years old, has a systolic blood pressure of 110 mm Hg, diastolic blood pressure of 70 mm Hg, and does not smoke? What is this person's corresponding probability of developing heart disease?

```{r q4-1}
x_age <- 45
x_ncigs <- 0
x_sbp <- 110
x_dbp <- 70

linearPart <- -9.12 + (0.0666 * x_age) + (0.0242 * x_ncigs) + (0.019 * x_sbp) + (0.008 * x_dbp)
linearPart # this is log odds

estProb <- exp(linearPart) / ( 1 + exp(linearPart))
estProb 

estProb / (1 - estProb) # this is probability

# try predict
newdata<-data.frame(age=45, sbp=110, dbp=70, ncigs=0)
predict(result, newdata) #is log odds

predict(result, newdata, type="response") # this is probability
 

```

## 5. Carry out the relevant hypothesis test to check if this logistic regression model with ...
the four predictors is useful in estimating the odds of heart disease. Clearly state the null and alternative hypotheses, test statistic, and conclusion in context.

H0: All coefs are 0
Ha: at least one coef is not equal to 0.

**Deviance Delta, aka delta G^2 == null deviance - residual deviance**

```{r q5-1}
deltaG2 <- result$null.deviance - result$deviance
deltaG2

1-pchisq(deltaG2,4)
```
**What is the test statistic? 111.2295????**

Reject H0, at least one coef is not 0.


## 6. Based on the Wald test, is diastolic blood pressure a significant predictor of heart disease, when the other three predictors are already in the model?

**Wald Test:** to remove a single predictor. Based on Z distribution. Analogous to t test in MLR

Bj-hat - 0 / se(Bj-hat)

H0: B for dbp is 0
Ha: B for dbp is not 0

dbp          0.007867   0.010057   z =0.782  0.43406

```{r q6-1}
zscore <- 0.007867  / 0.010057
zscore

(1 - pnorm(zscore)) * 2 # gives p value
```

## 7. Suppose a co-worker of yours suggests fitting a logistic regression model without the ...
two blood pressure variables. Carry out the relevant hypothesis test to check if this model without the blood pressure variables should be chosen over the previous model with all four predictors.

H0: coefs for sbp and dbp are 0
Ha: at least one of the sbp or dbp coefs are not 0.

```{r q7-1}

reduced <- glm(chd69 ~ age + ncigs, family = "binomial")

1-pchisq(reduced$deviance - result$deviance,2)

reduced$deviance - result$deviance

```
We reject the null, we need at least one of the bp predictors.


## 8. Based on all the analysis performed, which of these four predictors would you use in your logistic regression model?


age, ncigs, and sbp






























