---
title: "Module10HW"
author: "Diana McSpadden"
date: "11/4/2020"
output:
  word_document: default
  html_document: default
---

# Stat 6021: Homework Set 9

## H. Diana McSpadden
## UID: hdm52
## Date: 11/06/2020

**Attended group with**: Wright, Nam, Barbre


## Question 1. 
For this question, you will use a data set Boston, which comes from the MASS package in R. This is the same data set that you saw in the live session. The purpose of this question is to classify a town as being a high- or low-crime town based on some predictors. The governor of Massachusetts is most interested in identifying towns that have a high crime rate.


### Question 1 (a) Before fitting your logistic regression model ...
you will need to create a new variable. The variable crim is the per capita crime rate of the town. Create a new variable that classifies crim in the following manner: 

Define a town to have a:
* low crime rate, if its crime rate is less than 1 per capita.
* high crime rate, otherwise

Using the ifelse() function will be very helpful for this.

I highly recommend to add this newly created variable to the existing data frame.

Also, use the contrasts() function to see which class is the reference class for this newly created variable. What is the reference class?

**Work on Q1a**

```{r q1-setup1}
library(MASS)
library(Hmisc)
library(dplyr)
library(nnet)
library(ROCR)
```

``` {r q1a-1}

#?Boston


# even though we dont need it, here is the summary
aboutBoston <- summary(Boston$crim)
aboutBoston

Boston <- Boston %>%
     mutate(crimHigh = (crim >= 1.0) * 1)
```

```{r q1a-2}

Boston

attach(Boston)

```
```{r q1a-3}
is.factor(crimHigh)

##tell R to treat this variable as categorical
crimHigh<-factor(crimHigh)

is.factor(crimHigh)

levels(crimHigh)<-c("low","high")
levels(crimHigh)

contrasts(crimHigh)

```
**Answer Q1a**:

The reference class is low == 0

### Question 1 (b) Randomly split the data into a testing and training set of equal sizes. 
For consistency of results among all groups, use set.seed(199). Next, using the training set, fit a logistic regression model, with your newly created variable as the binary response variable, and with the following predictors: indus, nox, rad, tax, lstat, and medv. Then validate your model on the test data by creating an ROC curve.


What does your ROC curve tell you?

**Work on Q1b**

First, split the dataset into training and test:
```{r q1b-1}
detach(Boston) #detach Boston so variable names aren't a mess

set.seed(199) #set seed

sample<-sample.int(nrow(Boston), floor(.50*nrow(Boston)), replace = F) # get random sample to rows
train<-Boston[sample, ]
test<-Boston[-sample, ]
```

Second, fit the model on training data with specified predictors:

```{r q1b-2}
result<-glm(crimHigh ~ indus + nox + rad + tax + lstat + medv, family = "binomial", data=train)
```

Third, create ROC curve using validation, i.e. test, data.
```{r q1b-3}
##predicted prob of high crime vs. low crime rate for testing data based on training data
preds<-predict(result,newdata=test, type="response") # need to use type=response for probabilities.

##produce the numbers associated with classification table
rates<-prediction(preds, test$crimHigh) #did it correctly identify CHD diagnosis?

##store the true positive and false postive rates
roc_result<-performance(rates,measure="tpr", x.measure="fpr")
```


Now, plot the resulting ROC:
``` {r q1b-4}
##plot ROC curve and overlay the diagonal line for random guessing
plot(roc_result, main="ROC Curve for High Crime")
lines(x = c(0,1), y = c(0,1), col="red")
lines(x = c(0.08,0.08), y = c(0,1), col="blue")
lines(x = c(0,1), y = c(0.93,0.93), col="blue")

```
**Answer Q1b**:

I really am impressed with the predictive power of the model to predict high crime based on indus, nox, rad, tax, lstat, and medv.

This ROC curve shows there is some unknown threshold that predicts with a true positive rate around 93% with a corresponding false positive rate of only ~8%.

### Question 1(c) Find the AUC associated with your ROC curve. 
What does your AUC tell you?

**Work on Q1c**

Calculate the AUC...
```{r q1c-1}
##compute the AUC
aucvalues <- performance(rates, measure = "auc")
auc <- aucvalues@y.values[[1]]
auc
```
**Answer Q1c**:
The AUC = 0.978 means our model can perform with that accuracy when making a classification.

### Quest 1(d) Create a confusion matrix using a cutoff of 0.5. 
What is the false positive rate?

What is the false negative rate? 

Note: Be very careful with the coding associated with your response variable. 

You may want to use the levels() function to check how R was coding your response variable.

**Work on Q1d**

Create the confusion matrix:
```{r q1d-1}
table(test$crimHigh, preds>0.5)
```
0 == low crime

1 == high crime

```{r q1d-2}

trueNeg <- 164
falseNeg <- 11
truePos <- 75
falsePos <- 3

# here are all my calculations
overallErrorRate <- (falseNeg + falsePos) / (falseNeg + falsePos + trueNeg + truePos)
print(paste("Overall Error Rate: ",overallErrorRate))

falsePosRate <- falsePos / (trueNeg + falsePos)
print(paste("False Positive Rate: ",falsePosRate))


falseNegRate <- falseNeg / (falseNeg + truePos)
print(paste("False Negative Rate: ",falseNegRate))

#sensitivity <- 1 - falseNegRate
#sensitivity

#specifiity <- 1 - falsePosRate
#specifiity

```

**Answer Q1d**

False Positive Rate:  0.018

False Negative Rate:  0.128

### Question 1(e) Bearing in mind ...
the governor is most interested in identifying towns with high crime rates, how would you adjust the cutoff value from 0.5? Briefly explain why.

**Answer Q1e**

I am interpreting "most interested" as wanting to make sure as many high crime areas are identified as feasible with the model.

We could set the threshold lower than 0.5 to identify more high crime towns, i.e. fewer false negatives. We would however also be increasing the false positive rate.

## 2. (No R required) ...
A study was undertaken to determine the association between several predictors and the duration of pregnancies. 

The response variable, pregnancy duration, was recorded as a three-class variable: 
* pre-term for pregnancies lasting less than 36 weeks, 
* intermediate term for pregnancies lasting between 36 and 37 weeks, 
* and full term for pregnancies lasting more than 37 weeks. 

The predictors are:
* nutrition: index of nutritional status, higher scores denote better nutritional status
* less20 : 1 =less than 20 years old, 0 otherwise
* greater30 : 1 =greater than 30 years old, 0 otherwise
* alcohol: 1 =drank alcohol during pregnancy, 0 otherwise
* smoking: 1 =smoked during pregnancy, 0 otherwise


A first-order multinomial logistic regression is carried out for this study, and the R output is shown below.

Call:

multinom(formula = preg ~ nutrition + less20 + greater30 + alcohol + smoking)

...

### Quest 2(a) Write down the estimated multinomial logistic regression models associated with this analysis.

**Answer Q2a**

Reference class is **full term pregnancies**. There are 2 logits that are each compared against probability of full term pregnancy.

probability of preterm vs full term = exp(5.48 - (0.065 * nutrition) + (2.96 * less20) + (2.06 * greater30) + (2.04 * alcohol) + (2.45 * smoking))

probability of intermediate vs full term = exp(3.96 - (0.046 * nutrition) + (2.91 * less20) + (1.89 * greater30) + (1.07 * alcohol) +(2.23 * smoking))


### Question 2 (b) Calculate the Wald test statistics associated with the predictor alcohol, ...
and find the corresponding p-value. 

What are the conclusions in context at significance level α = 0.05? 
You do not need to apply the Bonferroni method here.

**Work on Q2b**

Calculate z and p values for the coefficients:
```{r q2c-1}

#preTermNutritionZ <- -0.065 / 0.018
#preTermNutritionP <- (1 - pnorm(abs(preTermNutritionZ)))*2
#print(paste("Preterm Nutrition Z Score: ", preTermNutritionZ))
#print(paste("Preterm Nutrition P Value: ", preTermNutritionP))

#preTermLess20Z <- 2.96 / .096
#preTermLess20P <- (1 - pnorm(abs(preTermLess20Z)))*2
#print(paste("Preterm Less 20 Z Score: ", preTermLess20Z))
#print(paste("Preterm Less 20 P Value: ", preTermLess20P))

#preTermGreater30Z <- 2.06 / 0.89
#preTermGreater30P <- (1 - pnorm(abs(preTermGreater30Z)))*2
#print(paste("Preterm Greater 30 Z Score: ", preTermGreater30Z))
#print(paste("Preterm Greater 30 P Value: ", preTermGreater30P))

preTermAlcoholZ <- 2.04 / 0.71
preTermAlcoholP <- (1 - pnorm(abs(preTermAlcoholZ)))*2
print(paste("Preterm Alcohol Z Score: ", preTermAlcoholZ))
print(paste("Preterm Alcohol P Value: ", preTermAlcoholP))

#preTermSmokingZ <- 2.45 / 0.73
#preTermSmokingP <- (1 - pnorm(abs(preTermSmokingZ)))*2
#print(paste("Preterm Smoking Z Score: ", preTermSmokingZ))
#print(paste("Preterm Smoking P Value: ", preTermSmokingP))

#intNutritionZ <- -0.046 / 0.015
#intNutritionP <- (1 - pnorm(abs(intNutritionZ)))*2
#print(paste("Intermediate Nutrition Z Score: ", intNutritionZ))
#print(paste("Intermediate Nutrition P Value: ", intNutritionP))

#intLess20Z <- 2.91 / 0.86
#intLess20P <- (1 - pnorm(abs(intLess20Z)))*2
#print(paste("Intermediate Less 20 Z Score: ", intLess20Z))
#print(paste("Intermediate Less 20 P Value: ", intLess20P))

#intGreater30Z <- 1.89 / 0.81
#intGreater30P <- (1 - pnorm(abs(intGreater30Z)))*2
#print(paste("Intermediate Greater 30 Z Score: ", intGreater30Z))
#print(paste("Intermediate Greater 30 P Value: ", intGreater30P))

intAlcoholZ <- 1.07 / 0.65
intAlcoholP <- (1 - pnorm(abs(intAlcoholZ)))*2
print(paste("Intermediate Alcohol Z Score: ", intAlcoholZ))
print(paste("Intermediate Alcohol P Value: ", intAlcoholP))

#intSmokingZ <- 2.23 / 0.67
#intSmokingP <- (1 - pnorm(abs(intSmokingZ)))*2
#print(paste("Intermediate Smoking Z Score: ", intSmokingZ))
#print(paste("Intermediate Smoking P Value: ", intSmokingP))
```

**Answer Q2b**:
Alcohol is significant when predicting preterm vs full term, but not when predicting intermediate term vs full term in this model.

For preterm vs fullterm, because the p-value is < 0.05, we reject H0, and say alcohol is significant in this model.

For intermediate vs. fullterm, because the p-value is < 0.05, we fail to reject H0, and say alcohol is not significant in this model.


### Question 2(c) Calculate the 95% confidence intervals associated with the predictor alcohol, ...
and interpret these intervals in context, in terms of relative risk of having a pregnancy that is pre-term, intermediate, or full term. You do not need to apply the Bonferroni method here.

**Work on 2c**

**Answer 2C**

```{r q2c-2}

preBhat <- 2.0429
preBse <- 0.7097461

preciLow <- preBhat - (1.96 * preBse)
preciHigh <- preBhat + (1.96 * preBse)

print(paste("preterm vs full term alcohol CI: ",preciLow, " - ",preciHigh))
print(paste("EXP preterm vs full term alcohol CI: ",exp(preciLow), " - ",exp(preciHigh)))


intBhat <- 1.067001
intBse <- 0.6495262
  
intciLow <- intBhat - (1.96 * intBse)
intciHigh <- intBhat + (1.96 * intBse)

print(paste("intermediate vs full term alcohol CI: ",intciLow, " - ",intciHigh))
print(paste("EXP intermediate vs full term alcohol CI: ",exp(intciLow), " - ",exp(intciHigh)))
```

**Answer Q2c**
95% CI for preterm vs full term coefficient within our model is: preterm 0.652 - 3.43. We are 95% confident the relative risk of preterm vs full term pregnancies when alcohol is consumed, and all other variables are held constant ,is between 1.92 and 31.

The 95% CI for the alcohol coefficient is between -0.21 and 2.34; since 0 is within the 95% CI, this CI agrees with our Wald's test that when predicting intermediate pregnancy length vs full length pregnancies alcohol is not a significant predictor within our model.







